Untitled

### 2024-03-04 ms/ms.qmd
working on ms

In preamble:
- short codes for commonly used symbols

@@

-   the native light habitat or environment

@@

@@

We distinguished among these hypotheses by comparing AA among wild tomato species from different native light habitats, grown under simulated sun and shade, and measured under contrasting light intenstiy (Figure of hypotheses and predictions). We measured AA on XXX individual plants from XXX accessions (average of X replicates per light treatment) using a recently developed method [@triplett_amphistomy_2024]. With this method, we directly compare the photosynthetic rate of an untreated amphistomatous leaf to that of the same leaf with gas exchange blocked through the adaxial (upper) surface by transparent plastic, which we refer to as 'pseudohypostomy'. To compare amphi- and pseudohypostomatous leaves at identical whole-leaf $g_\text{sc}$, we measure $A$ over a range of $g_\text{sc}$, inducing stomatal opening and closure by modulating humidity (see Materials and Methods for further details).

@@

>>>>

```{r load-objects, echo=FALSE, eval=TRUE}

# I am mostly putting placeholders here for objects/values that will need to be
# loaded eventually to use in paper.
n_plant = 600
n_acc = 30
mean_n_rep = 10 # average number of individual replicates for accession per light treatment

# environmental condition in growth room
```

## Notes

In preamble:
- short codes for commonly used symbols
- ?

@@

-   the native light habitat or environment: the SPLASH paper refers to what they calculated as 'habitat' PPFD, so I am going to use that terminology

@@

## Main

@@

We distinguished among these hypotheses by comparing AA among wild tomato species from different native light habitats, grown under simulated sun and shade light treatments, and measured under contrasting light intensity (Figure of hypotheses and predictions). We measured AA on `r n_plant` individual plants from `r n_acc` accessions (average of `r mean_n_rep` replicates per light treatment) using a recently developed method [@triplett_amphistomy_2024]. With this method, we directly compare the photosynthetic rate of an untreated amphistomatous leaf to that of the same leaf with gas exchange blocked through the adaxial (upper) surface by transparent plastic, which we refer to as 'pseudohypostomy'. To compare amphi- and pseudohypostomatous leaves at identical whole-leaf $g_\text{sc}$, we measure $A$ over a range of $g_\text{sc}$, inducing stomatal opening and closure by modulating humidity (see Materials and Methods for further details).

Table of directional predictions (with table summarizing results? part of conceptual figure? in supplement?)

caption: Directional predictions associated with each hypothesis explaining why amphistomy advantage (AA) might be greater for leaves in sunny, open habitats. For each hypothesis, we make predictions for how native light habitat, light treatment, and light intensity would affect AA.

| hypothesis | native light habitat | light treatment | light intensity |
|:----------:|:--------------------:|:---------------:|:---------------:|
| acclimatory | cor(PAR, AA) = 0 | AAsun = AAshade | AA2000 > AA150 |
| plastic | cor(PAR, AAsun) > 0 | AAsun > AAshade | AA2000 = AA150 |
| constitutive | cor(PAR, AA) > 0 | AAsun = AAshade | AA2000 = AA150 |

[add rows for when multiple hypotheses are supported simultaneously? or put that in SI?]

@@

## Materials and Methods

[this will be moved to SI eventually]

<!-- work until 10:15, then email Sarah about scheduling meeting -->

### Accessions

We compared AA among `r n_acc` ecologically diverse accessions of wild tomato, including representatives of all described species of *Solanum* sect. *Lycopersicon* and sect. *Lycopersicoides* [@peralta_taxonomy_2008] and the cultivated tomato *S. lycopersicum* var. *lycopersicum* (tab:accessions). Seeds provided by the Tomato Genetics Resource Center germinated on moist paper in plastic boxes after soaking for 30-60 minutes in a 50\% (volume per volume) solution of household bleach and water, followed by a thorough rinse. We transferred seedlings to cell-pack flats containing Pro-Mix BX potting mix (Premier Tech, Rivière-du-Loup, Quebec, Canada) once cotyledons fully emerged, typically within 1-2 weeks of sowing. After seedlings established in cell-pack flats for $\approx 2$ weeks, we transplanted them to 3.78 L plastic pots containing Pro-Mix BX potting mix, coral sand (Pro-Pak, Honolulu, Hawaiʻi, USA), and cinders (Niu Nursery, Honolulu, Hawaiʻi, USA) [look up exact proportions of each].

outline of tab:accessions
LA number,Species name, lat, lon, elev, SPLASH light value
